master thesis

### 2018-03-23 Paper - Latex/Paper - Skizze - v4.tex
Paper - Tables with bold header; more data; Binomial regression; Simulation results

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{real data analysis}
    \todo{Evaluation section in methods?}
    \begin{itemize}
      \item Import tree
      \item Import interactions
      \item run castor algorithm / and others?
      \item interprete results (cross validation - leave 100 out)
    \end{itemize}
    We then evaluated the results in two ways. On the one hand, we have examined the results of some 
      sub-trees for their credibility with the help of biological background knowledge. On the other 
      hand, we performed a leaf 100 out cross validation.

@@

  We then evaluated the results in two ways. On the one hand, we have examined the results of some 
    sub-trees for their credibility with the help of biological background knowledge (Subsection 
    \ref{sec:biological view}). On the other hand, we performed a leaf 100 out cross validation
    (Subsection \ref{sec:results cross evaluation}).

@@

    A tree is multifurcated if there are nodes that have more than two children. In the case of a 
      phylogenetic tree, the ridge of multifurcation describes the amount of lack of information about 
      the topology of the tree. A complete phylogenetic tree would be binary, meaning that the number 
      of leaf nodes equals the number of internal nodes minus one. Since we only work with a 
      artificially constructed tree (synthesis of several phylogenetic trees), this tree is strongly 
      multifurcated: It has 241,974 internal nodes for 2,293,463 leaf nodes $\rightarrow$ 98.16 \% 
      missing nodes. \\
    We calculated these percentages of missing nodes and also missing state information for some 
      subtrees and plotted them in table \ref{table:percentage loss information subtrees}. \\
    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l||r|r| }
          \hline
          Subtree of & unknown states & multifurcation \\ 
          \hline \hline
          Eukaryota       & 97.34~\%  & 98.16~\% \\
          \hline \hline
          Metazoa         & 96.44~\%  & 87.93~\% \\ \hline
          Fungi           & 98.87~\%  & 96.97~\% \\ \hline
          Chloroplastida  & 99.14~\%  & 89.46~\% \\
          \hline \hline            
          % Arthropoda      & 97.49~\%  & 89.95~\% \\ \hline
          Apicomplexa     & 86.26~\%  & 87.16~\% \\ \hline
          Nematoda        & 89.01~\%  & 88.59~\% \\ \hline
          Chordata        & 88.59~\%  & \cellcolor{green!50}66.49~\% \\ \hline
          Platyhelminthes & \cellcolor{green!50}68.73~\%  & 80.34~\% \\
          \hline \hline            
          Insecta         & 97.11~\%  & 90.78~\% \\
          \hline  
        \end{tabular}
      \end{center}
      \caption{Examination of subtrees regarding missing information. \\ 
        The percentage values show the ratio of missing information of: unknown states (missing state 
          information of leaf nodes) and multifurcation (missing internal nodes). \\
        The subtrees are from different taxa: domain (Eukaryota), kingdom (Metazoa, Fungi, 
          Chloroplastida), phylum (Apicomplexa, Nematoda, Chordata, Platyhelminthes) and class 
          (Insecta). \\
        The two by far smallest values were highlighted in green.} 
      \label{table:percentage loss information subtrees} 
    \end{table}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Poisson regression of the multifurcation}
      For a regression analysis of the multifurcation first set up a simple generalized linear model and 
        then output the intercept of it:
      \lstset{language=R}
      \begin{lstlisting}[gobble=6]
        kingdom.furc.mod <- glm(multifurc~kingdom, 
                                    data=inner.taxa,
                                        family="poisson")
        # get the intercept:
        summary(kingdom.furc.mod)
      \end{lstlisting}
      Since the multifurcation can also be 0, the intercept describes the expected mean value and thus, 
        if it is greater than 0, it says that we have a multifurcation: $2.62 > 0$ $\Rightarrow$ there 
        is a multifurcation. \\
      We next compared the intercept of the different kingdoms and found that
        $$1.22 \text{ (Fungi intercept)} > -0.11 \text{ (Metazoa intercept)} > -0.09 \text{ (Chloroplastida intercept)}$$
        % Coefficients:
        %                                         Estimate Std. Error z value Pr(>|z|)
        % (Intercept)                             2.624897   0.003641  721.00   <2e-16
        % kingdom['ott352914', 'Fungi']           1.224650   0.004054  302.10   <2e-16
        % kingdom['ott361838', 'Chloroplastida'] -0.090993   0.003960  -22.98   <2e-16
        % kingdom['ott691846', 'Metazoa']        -0.114873   0.003732  -30.78   <2e-16
        that means the strength of the multifurcation \colorbox{red}{is the same/has the same size 
        sorting}. \todo{kann das sein?} \\
    %---------------------------------------------------------------------------------------------------

      We investigated several models that could describe multifurcation: \\
      In doing so, we allowed the different influence of the taxa and the heights and depths of a node 
        to be included. From this we got 9 times 4 models of different complexity levels (see first row 
        of table \ref{table:BIC multifurcation}). \\
      We first compared the models based on their residuals. These give us the error of the model. If 
        the value is small, our data will be well modeled. These values are meaningful only for the same 
        complexity, therefore we then calculated the BIC (Bayesian information criterion) values of the 
        models and used them to evaluate the models.
      \lstset{language=R}
      \begin{lstlisting}[gobble=6]
        anova(kingdom.furc.mod, phylum.furc.mod, class.furc.mod, 
                      order.furc.mod, family.furc.mod, test="LRT")
        BIC(kingdom.furc.mod)
        ...
      \end{lstlisting}
      The deviance table with the residuals can be found in the appendix (table 
        \ref{table:Residuals multifurcation}). Table \ref{table:BIC multifurcation}. lists the BICs.

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            Model / Taxa & Kingdom & Phylum & Class & Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models \\
          These models were created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC multifurcation} 
      \end{table}
      % * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
      Within the every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Smaller taxa than orders (eg 
        family) were computationally too expensive to calculate.

@@

      Next to the problem of the multifurcation of the tree is the less of data we have for the species.
        For the ancestral state reconstruction, we need information in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5 346 414 interactions 
        (at this timepoint). Out of this data we got 51,337 distinct free-living species and 47 332 
        distinct parasite species $\rightarrow$ unknown nodes 2,194,794 ($\approx 95.7~\%$). \\
      We found also 57,352 (not distinct!) source species and 809,993 (not distinct!) target
        species without OTT ids. Since we currently use only OTT ids, we could not use this information. \\
      % \todo{With this only $\approx 4.3~\%$ information in our leaf nodes are ...} \\

@@

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            Model / Taxa & Kingdom & Phylum & Class & Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BICs of unknown information}
        \label{table:BIC unknown information} 
      \end{table}
      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better. \\
      Again, the calculation of finer models (eg order or family) was too expensive. \\

@@

    As presented, we compare two methods in our simulation to their prediction accuracy: Fitch and 
      Sankoff. \\
    We examine different parameters. In figure \ref{fig:influence of unknown data} is an overview of 
      the results. \\
    The first column describes the distributions of free-livings and parasites with a given threshold 
      for the respective simulations to the right. \\
    The middle column investigates the influence of the unknown states, the right the influence of the 
      strength of the multifurcation. \\
    The y-axes indicate the percentage of correctly predicted states (including known states). On the 
      x-axis the percentage of forgotten states or missing internal nodes. \\
    Each point corresponds to the average of one hundred simulations, each with 10,000 leaf nodes. \\
    For the middle column we set the strength of the multifurcations to 0.95~\% similar to the real 
      data and in the right column the amount of the unknowns to 0.95~\% also similar to the real data. \\

    As \colorbox{red}{you} can see both algortithms are always over 50~\% and therefore better than guessing. Moreover, 
      they are usually close to each other, with Sankoff always makes better predictions except for 
      equally distributed states as Fitch. \\
    \colorbox{red}{You} can see that the distribution of states has a strong influence on the prediction. The more 
      evenly distributed, the harder it is to predict. \\
    If more than 60~\% of internal nodes are missing, Fitch breaks significantly in his prediction 
      compared to Sankoff.

    \begin{figure}
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
      \caption{Influence of unknown data to prediction}
      \label{fig:influence of unknown data}
    \end{figure}

    In the end, Sankoff is in most cases the more accurate algortihmus and was therefore used for our 
      prediction of the real data.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of the real data analysis created with Sankoff}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Biological view} \label{sec:biological view}
      To analyze the results, we have selected some phyla (sub-trees) to evaluate our results 
        selectively from the biological point of view: Chordata, Nematoda, Platyhelminthes and 
        Apicomplexa. \\
      In Table \ref{table:phylum leaf nodes states} we compare the given states with the predicted ones. \\
      There are several factors that play a role here, and in part, crystallize in these examples. \\
      An important factor here is that the credibility of the results with the accuracy of the input 
        data, i. the data of GloBI, stands and falls. Errors of incorrect input data can be amplified by 
        incorrect prediction of unknown species and can be reversed in order to improve the data 
        situation of GloBI. \\
      Since we look at such large trees we can not expect to know all the parasites, so we look at 
        individual positives. Positive in the sense that the majority have the opposite state.

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r| }
            \hline
            & & \multicolumn{2}{c||}{original states} & \multicolumn{4}{c|}{final states} \\
            Phylum & \# nodes       & FL & P              & 0 (FL) & 0.3 & 0.5 & 1 (P) \\
            \hline \hline
            Chordata & 91785        & 10451 & 18          & 91759 & 0 & 0 & 26 \\
            &                       & 99.83~\% & 0.49~\%  & 99.97~\% & & & 0.03~\% \\ \hline
            Nematoda & 30127        & 21 & 3289           & 1604 & 142 & 1196 & 27185 \\
            &                       & 0.63~\% & 99.37~\%  & 5.32~\% & 0.47~\% & 3.97~\% & 90.23~\% \\ \hline
            Platyhelminthes & 22683 & 7 & 7086            & 175 & & & 22508 \\
            &                       & 0.1~\% & 99.9~\%    & 0.77~\% & 0 & 0 & 99.23~\% \\ \hline
            Apicomplexa & 1863      & 1 & 255             & 1 & 0 & 0 & 1862 \\
            &                       & 0.39~\% & 99.61~\%  & 0.05~\% & & & 99.95~\% \\
            % \hline \hline
            % Phylum & \# nodes & FL & P
            %   & 0 (FL) & 0.4 & 0.5 & 0.67 & 0.75 & 1 (P) \\
            % Arthropoda & 1198981 & 18912 & 11141 
            %   & 1099509 & 1313 & 22478 & 4176 & 1665 & 70223 \\
            % & & 62.93~\% & 37.07~\%
            %   & 91.7~\% & 0.11~\% & 1.87~\% & 0.35~\% & 0.14~\% & 5.86~\% \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{Phylum (leaf nodes)}
        \label{table:phylum leaf nodes states} 
      \end{table}

@@

      Most species of Platyhelminthes (flatworms) are parasites, although there are also free-living, 
        predatory feeding species. These are summarized in the Turbellaria, while the parasites are 
        divided into three other classes \cite{Ax1961}. This also corresponds to our observations. There 
        is one class (Rhabditophora) that contains all but one single exception of free-living species 
        of this phylum, which includes the Turbellaria. \\
      It should be noted, however, that this classification is outdated, as it has been proven that 
        the Turbellaria are not monophyletic. But we will not go into that here. \\
      For the Platyhelminthes we had more state information for the leaf nodes compared to the other 
        considered subtrees, see Table \ref{table:percentage loss information subtrees}. We start with 
        0.1~\% free-livings and predicted 0.77~\% as free-living spezies. \\
    %---------------------------------------------------------------------------------------------------

@@

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r| }
            \hline
            & & \multicolumn{2}{c||}{original states} & \multicolumn{6}{c|}{final states} \\
            Phylum & \# nodes & FL & P
              & 0 (FL) & 0.4 & 0.5 & 0.67 & 0.75 & 1 (P) \\
            \hline \hline
            Chordata & 122546 & 10451 & 18 
              & 122473 & 0 & 0 & 0 & 0 & 73 \\
            Nematoda & 33564 & 21 & 3289 
              & 846 & 0 & 1133 & 0 & 0 & 31585 \\
            Platyhelminthes & 27142 & 7 & 7086 
              & 1010 & 0 & 175 & 0 & 0 & 25957 \\
            Apicomplexa & 2102 & 1 & 255 
              & 1 & 0 & 0 & 0 & 0 & 2101 \\
            % \hline
            % Arthropoda & 1319460 & 18912 & 11141 
            %   & 1207204 & 1313 & 25499 & 4852 & 1957 & 78635 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Phylum (inkl internal nodes)}
        \label{table:Phylum internal nodes}
      \end{table}

>>>>

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{real data analysis}
    \todo{Evaluation section in methods?}
    \begin{itemize}
      \item Import tree
      \item Import interactions
      \item run castor algorithm / and others?
      \item interprete results (cross validation - leave 100 out)
    \end{itemize}
    We then evaluated the results in two ways. On the one hand, we have examined the results of some 
      subtrees for their credibility with the help of biological background knowledge. On the other 
      hand, we performed a leaf 100 out cross validation.

@@

  We then evaluated the results in two ways. On the one hand, we have examined the results of some 
    subtrees for their credibility with the help of biological background knowledge (subsection 
    \ref{sec:biological view}). On the other hand, we performed a leaf 100 out cross validation
    (Subsection \ref{sec:results cross evaluation}).

@@

    A tree is multifurcated if there are nodes that have more than two children. In the case of a 
      phylogenetic tree, the ridge of multifurcation describes the amount of lack of information about 
      the topology of the tree. A complete phylogenetic tree would be binary, meaning that the number 
      of leaf nodes equals the number of internal nodes minus one. Since we only work with a 
      artificially constructed tree (synthesis of several phylogenetic trees), this tree is strongly 
      multifurcated: It has 241,974 internal nodes for 2,293,463 leaf nodes $\rightarrow$ 98.16 \% 
      missing nodes. \\
    We calculated these percentages of missing nodes and also missing state information for some 
      subtrees and plotted them in table \ref{table:percentage loss information subtrees}. \\
    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l||r|r| }
          \hline
          \bfseries Subtree of & \bfseries Unknown States & \bfseries Multifurcation \\ 
          \hline \hline
          Eukaryota       & 97.34~\%  & 98.16~\% \\
          \hline \hline
          Metazoa         & 96.44~\%  & 87.93~\% \\ \hline
          Fungi           & 98.87~\%  & 96.97~\% \\ \hline
          Chloroplastida  & 99.14~\%  & 89.46~\% \\
          \hline \hline            
          % Arthropoda      & 97.49~\%  & 89.95~\% \\ \hline
          Apicomplexa     & 86.26~\%  & 87.16~\% \\ \hline
          Nematoda        & 89.01~\%  & 88.59~\% \\ \hline
          Chordata        & 88.59~\%  & \cellcolor{green!50}66.49~\% \\ \hline
          Platyhelminthes & \cellcolor{green!50}68.73~\%  & 80.34~\% \\
          \hline \hline            
          Insecta         & 97.11~\%  & 90.78~\% \\
          \hline  
        \end{tabular}
      \end{center}
      \caption{Examination of subtrees regarding missing information. \\ 
        The percentage values show the ratio of missing information of: unknown states (missing state 
          information of leaf nodes) and multifurcation (missing internal nodes). \\
        The subtrees are from different taxa: domain (Eukaryota), kingdom (Metazoa, Fungi, 
          Chloroplastida), phylum (Apicomplexa, Nematoda, Chordata, Platyhelminthes) and class 
          (Insecta). \\
        The two by far smallest values were highlighted in green.} 
      \label{table:percentage loss information subtrees} 
    \end{table}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Poisson regression of the multifurcation}
      For a regression analysis of the multifurcation first set up a simple generalized linear model and 
        then output the intercept of it:
      \lstset{language=R}
      \begin{lstlisting}[gobble=6]
        kingdom.furc.mod <- glm(multifurc~kingdom, 
                                    data=inner.taxa,
                                        family="poisson")
        # get the intercept:
        summary(kingdom.furc.mod)
      \end{lstlisting}
      Since the multifurcation can also be 0, the intercept describes the expected mean value and thus, 
        if it is greater than 0, it says that we have a multifurcation: $2.62 > 0$ $\Rightarrow$ there 
        is a multifurcation. \\
      We next compared the intercept of the different kingdoms and found that
        $$1.22 \text{ (Fungi intercept)} > -0.11 \text{ (Metazoa intercept)} > -0.09 \text{ (Chloroplastida intercept)}$$
        % Coefficients:
        %                                         Estimate Std. Error z value Pr(>|z|)
        % (Intercept)                             2.624897   0.003641  721.00   <2e-16
        % kingdom['ott352914', 'Fungi']           1.224650   0.004054  302.10   <2e-16
        % kingdom['ott361838', 'Chloroplastida'] -0.090993   0.003960  -22.98   <2e-16
        % kingdom['ott691846', 'Metazoa']        -0.114873   0.003732  -30.78   <2e-16
        that means the strength of the multifurcation \colorbox{red}{is the same/has the same size 
        sorting}. \todo{kann das sein?} \\

      We investigated several models that could describe multifurcation: \\
      In doing so, we allowed the different influence of the taxa and the heights and depths of a node 
        to be included. From this we got 9 times 4 models of different complexity levels (see first row 
        of table \ref{table:BIC multifurcation}). \\
      We first compared the models based on their residuals. These give us the error of the model. If 
        the value is small, our data will be well modeled. These values are meaningful only for the same 
        complexity, therefore we then calculated the BIC (Bayesian information criterion) values of the 
        models and used them to evaluate the models. \\
      % \lstset{language=R}
      % \begin{lstlisting}[gobble=6]
      %   anova(kingdom.furc.mod, phylum.furc.mod, class.furc.mod, 
      %                 order.furc.mod, family.furc.mod, test="LRT")
      %   BIC(kingdom.furc.mod)
      %   ...
      % \end{lstlisting}
      The deviance table with the residuals can be found in the appendix (table 
        \ref{table:Residuals multifurcation}). Table \ref{table:BIC multifurcation}. lists the BICs.

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models were created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC multifurcation} 
      \end{table}
      % * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
      Within the every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (eg 
        family) were computationally too expensive to calculate. \\
      The model multifurc $\sim$ order * mean.height turns out to be the best of our models, whereby it 
        is possible that eg multifurc $\sim$ family * mean.height is better.

@@

      Next to the problem of the multifurcation of the tree is the less of interaction data we have for 
        the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 distinct free-living species and 47 332 
        distinct parasite species $\rightarrow$ unknown nodes 2,194,794 ($\approx 95.7~\%$). \\
      We found also 57,352 (not distinct!) source species and 809,993 (not distinct!) target
        species without OTT ids. Since we currently use only OTT ids, we could not use this information. \\
      % \todo{With this only $\approx 4.3~\%$ information in our leaf nodes are ...} \\
      \todo{discussion!!!} \\

@@

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BICs of unknown information}
        \label{table:BIC unknown information} 
      \end{table}
      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better. \\
      Again, the calculation of finer models (eg order or family) was too expensive. \\

@@

    As presented, we compare two methods in our simulation to their prediction accuracy: Fitch and 
      Sankoff. \\
    Figure \ref{fig:influence of unknown data} shows the results of examining various parameters.
    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 100mm 0mm 0mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
    \end{figure}
    ~ \\ ~ \\ ~ \\
    \begin{figure}
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 150mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
      \caption{Influence of unknown data to prediction. \\
        The first column describes the distributions of free-livings and parasites with a given 
          threshold for the respective simulations to the right. \\
        The middle column investigates the influence of the unknown states, the right the influence of
          the strength of the multifurcation. \\
        The y-axes indicate the percentage of correctly predicted states (including known states). On 
          the x-axis the percentage of forgotten states or missing internal nodes. \\
        Each point corresponds to the average of one hundred simulations, each with 10,000 leaf nodes. \\
        For the middle column we set the strength of the multifurcations to 0.95~\% similar to the 
          real data and in the right column the amount of the unknowns to 0.95~\% also similar to the 
          real data.}
      \label{fig:influence of unknown data}
    \end{figure}
    First, we tested different trasition probability distributions including threshold (first column). 
      It can be observed that the more balanced the percentages of free-livings and parasites, the 
      worse the prediction of the algorithms (second and third column). \\
    It can be seen that both algortithms are always over 50~\% prediction and therefore better than 
      guessing. Moreover, they are usually close to each other, with Sankoff always makes better 
      predictions except for equally distributed states as Fitch. \\
    On the other hand, we examined the influence of missing internal nodes (ridge of multifurcation) 
      and missing leaf node information (unknown leaf nodes). \\
    Both factors have a relatively linear influence on the Sankoff method. Fitch, on the other hand, 
      breaks significantly in his prediction from about 70\% unknown leaf nodes or mutifurcation. \\

    Since we have a lot of missing data in most subtrees (both internal nodes and state informations), 
      the Sankoff gives a better prediction and was thus used for real data analysis.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of the real data analysis created with Sankoff}
    This section is about evaluating the prediction of real data using the Sankoff method. \\
    It is divided into two subsections. \ref{sec:biological view}: The analysis of some subtrees using 
      biological background knowledge. \ref{sec:results cross evaluation}: Presentation of the results 
      of the leaf 100 out cross validation.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Biological view} \label{sec:biological view}
      To analyze the results, we have selected some phyla (subtrees) to evaluate our results 
        selectively from the biological point of view: Chordata, Nematoda, Platyhelminthes and 
        Apicomplexa. \\
      In Table \ref{table:phylum leaf nodes states} we compare the given states with the predicted ones. \\
      There are several factors that play a role here, and in part, crystallize in these examples. \\
      An important factor here is that the credibility of the results with the accuracy of the input 
        data, i. the data of GloBI, stands and falls. Errors of incorrect input data can be amplified by 
        incorrect prediction of unknown species and can be reversed in order to improve the data 
        situation of GloBI. \\
      Since we look at such large trees we can not expect to know all the parasites, so we look at 
        individual positives. Positive in the sense that the majority have the opposite state.

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r| }
            \hline
            & & \multicolumn{2}{c||}{\bfseries original states} & \multicolumn{4}{c|}{\bfseries final states} \\
            \bfseries Phylum & \bfseries \# nodes & \bfseries FL & \bfseries P & \bfseries 0 \bfseries (FL) & \bfseries 0.3 &\bfseries  0.5 & \bfseries 1 (P) \\
            \hline \hline
            Chordata & 91785        & 10451 & 18          & 91759 & 0 & 0 & 26 \\
            &                       & 99.83~\% & 0.49~\%  & 99.97~\% & & & 0.03~\% \\ \hline
            Nematoda & 30127        & 21 & 3289           & 1604 & 142 & 1196 & 27185 \\
            &                       & 0.63~\% & 99.37~\%  & 5.32~\% & 0.47~\% & 3.97~\% & 90.23~\% \\ \hline
            Platyhelminthes & 22683 & 7 & 7086            & 175 & & & 22508 \\
            &                       & 0.1~\% & 99.9~\%    & 0.77~\% & 0 & 0 & 99.23~\% \\ \hline
            Apicomplexa & 1863      & 1 & 255             & 1 & 0 & 0 & 1862 \\
            &                       & 0.39~\% & 99.61~\%  & 0.05~\% & & & 99.95~\% \\
            % \hline \hline
            % Phylum & \# nodes & FL & P
            %   & 0 (FL) & 0.4 & 0.5 & 0.67 & 0.75 & 1 (P) \\
            % Arthropoda & 1198981 & 18912 & 11141 
            %   & 1099509 & 1313 & 22478 & 4176 & 1665 & 70223 \\
            % & & 62.93~\% & 37.07~\%
            %   & 91.7~\% & 0.11~\% & 1.87~\% & 0.35~\% & 0.14~\% & 5.86~\% \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{Phylum (leaf nodes)}
        \label{table:phylum leaf nodes states} 
      \end{table}

@@

      Most species of Platyhelminthes (flatworms) are parasites, although there are also free-living, 
        predatory feeding species. These are summarized in the Turbellaria, while the parasites are 
        divided into three other classes \cite{Ax1961}. This also corresponds to our observations. There 
        is one class (Rhabditophora) that contains all but one single exception of free-living species 
        of this phylum, which includes the Turbellaria. \\
      It should be noted, however, that this classification is outdated, as it has been proven that 
        the Turbellaria are not monophyletic. But we will not go into that here. \\
      For the Platyhelminthes we had more state information for the leaf nodes compared to the other 
        considered subtrees, see Table \ref{table:percentage loss information subtrees}. We start with 
        0.1~\% free-livings and predicted 0.77~\% as free-living spezies. \\

@@

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r| }
            \hline
            & & \multicolumn{2}{c||}{\bfseries original states} & \multicolumn{6}{c|}{\bfseries final states} \\
            \bfseries Phylum & \bfseries \# nodes & \bfseries FL & \bfseries P
              & \bfseries 0 (FL) & \bfseries 0.4 & \bfseries 0.5 & \bfseries 0.67 & \bfseries 0.75 & \bfseries 1 (P) \\
            \hline \hline
            Chordata & 122546 & 10451 & 18 
              & 122473 & 0 & 0 & 0 & 0 & 73 \\
            Nematoda & 33564 & 21 & 3289 
              & 846 & 0 & 1133 & 0 & 0 & 31585 \\
            Platyhelminthes & 27142 & 7 & 7086 
              & 1010 & 0 & 175 & 0 & 0 & 25957 \\
            Apicomplexa & 2102 & 1 & 255 
              & 1 & 0 & 0 & 0 & 0 & 2101 \\
            % \hline
            % Arthropoda & 1319460 & 18912 & 11141 
            %   & 1207204 & 1313 & 25499 & 4852 & 1957 & 78635 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Phylum (inkl internal nodes)}
        \label{table:Phylum internal nodes}
      \end{table}
